Digital Twins for Ocean Robots

### 2024-11-12 paper/paper.tex
minor text edits / proofreading

DTOR is a framework to access and simulate the global fleet of ocean observing devices (or {\it ocean robots}) that monitor climate change. It brings these observations to Julia and let's us pair ocean robots with virtual counterparts (or {\it twins}). Digital twins provide a bridge to predictive models that enables machine learning. In turn, simulating observations in a virtual environment can help evaluate observational strategies a priori, during deployment, or afterwards. In this paper we present the framework, review simulated ocean robots, and discuss envisioned applications.

@@

The DTOR framework was introduced at the Symposium on Advances in Ocean Observation in 2022, and later presented to US-CLIVAR and JuliaCon in 2023. The primary goal of DTOR is to associate ocean observing systems (or {\it ocean robots}) with virtual counterparts (or {\it twins}), and do it for the whole fleet of ocean robots that are currently at sea or have observed the Ocean in the past (Fig.~\ref{fig:ocean-ops}). The virtual twins are to be created through numerical model simulations, which can come in different flavors and languages. The simulation of observing systems in the future, as climate change progresses, is also part of the scope of the DTOR project. At JuliaCon 2023, we presented a solution to (1) access and manipulate the data collected by ocean robots (inc. \pkg{OceanRobots.jl} and \pkg{ArgoData.jl}), and (2) simulate such observations in a virtual environment using a hierarchy of models (incl. \pkg{ClimateModels.jl} and \pkg{MITgcm.jl}).

The model hierarchy includes fast climate model emulators, global model output, ocean reanalyses, high-resolution models, and several ways to represent marine ecosystems. Through a streamlined workflow, \pkg{ClimateModels.jl} (JuliaCon 2021, 2023) makes it easy to operate these models that can provide a virtual environment for the virtual ocean robot fleet to observe and navigate. \pkg{Drifters.jl} \cite{Forget2021} can be use to predict pathways of ocean robots that tend to follow ocean currents. \pkg{MeshArrays.jl} adds basic geospatial support for global climate model grids, providing capabilities such as interpolation and geolocation on a grid. The framework leverages and links to a variety of highly capable Julia packages from the community. It notably provides extension for visualization with \pkg{Makie.jl}, and tutorial examples in the form of \pkg{Pluto.jl} notebooks.

@@

Part of the fleet of ocean robots currently at sea is depicted in Fig.~\ref{fig:ocean-ops}, focusing on some of the most common observing platforms. To create this map, \pkg{OceanRobots.jl} queries the meta-data-base from Ocean-OPS.org, which keeps track of the whole ocean fleet of scientific observation platforms in real time, and provides a restful API. \pkg{OceanRobots.jl} provides a simple interface to this API via the `OceanOPS` module. It then provides dedicated data structures to access and utilize data from common ocean robot types such as  `SurfaceDrifter` and `ArgoFloat` (shown in Fig.~\ref{fig:ocean-ops}). A non-exhaustive list of ocean observing platform categories within the scope of \pkg{OceanRobots.jl} is provided in Tab.~\ref{tab:robots}.

@@

Argo profiling floats (blue dots in Fig.~\ref{fig:ocean-ops}; Code~\ref{lst:ArgoCode}) are one of our main tools to monitor global warming below the sea surface. These devices provide a less detailed view of oceanic pathways (Fig.~\ref{fig:Argo_Float}, top panels) than surface drifters do (e.g., Fig.~\ref{fig:drifters}) with just one location being recorded every ten days. However, Argo floats have a crucial diving capability that surface drifters don't have -- they go up and down the water column to measure temperature and salinity (T,S; bottom panels of Fig.~\ref{fig:Argo_Float}). As a result, the advent of Argo in the early 2000s \cite{Roemmich2019} opened up a whole new era of geospatial analysis, state estimation, and parameter inference over the global Ocean  \cite{Forget2007,Forget2010,MFF14,Forget2015a,FFL15,FP15,Forget2024a}. There were 3837 Argo floats at sea on `2024/05/27`, and a total of 18730 in the Argo data base. Since the Argo array is such an important and somewhat complex observing system, a dedicated package called \pkg{ArgoData.jl} was created that \pkg{OceanRobots.jl} uses under the hood.

@@

\pkg{OceanRobots.jl} brings these key climate data sets to the Julia community. It provides a bridge to climate scientists working on observations and models, who are interested in leveraging the powerful Julia software ecosystem (e.g., for numerical modeling, machine learning, and statistical analysis), and have much expertise to contribute. The development of \pkg{OceanRobots.jl} aims to help advance (1) how we understand and simulate observations of climate change, and (2) climate literacy and education by providing simple apps that anyone should be able to use.

@@

Our ability to simulate the data produced by ocean robots is directly linked to our ability, or lack thereof, to explain observed variations and decipher mechanisms from data. Numerical modeling is thus often motivated and driven by observations. Many research activities involve combining ocean models and observations, with the typical goal of learning model parameters and dynamics from data \cite{Forget2010,Forget2015a,FFL15,FP15,Forget2024a}. Observing system simulations in a virtual environment (using models) enable a wide range of applications -- to test out innovative platform or sensor designs ahead of deployment, to optimize global monitoring strategies, or to guide deployed assets in real time for example.

The two simple examples presented below illustrate the simulation of (1) environmental sensors and (2) observing platforms. First, in Fig.~\ref{fig:Argo_Float_simu} we simulate an Argo data collection for temperature (T) and salinity (S) by sampling an ocean climatology along the track of the Argo float (the one from Fig.~\ref{fig:Argo_Float}). This calculation requires (1) a model prediction of $T,S(x,y,z,t)$, which can be based on statistical or mechanistic models, e.g. on a discrete grid, and (2) tools that deal with the Earth geography, can localize observations on a grid, and perform interpolation tasks.

@@

Our second example focuses on the simulation of observing platforms moving with ocean currents. This basic simulation of surface drifter pathways uses the geostatistical average of velocities calculated in Fig.~\ref{fig:drifters_ve} as a predictive model for the $U,V(x,y,t)$ velocity fields. This gridded data set is simply provided as input to \pkg{Drifters.jl}, which then releases virtual drifters and calculates their trajectories following the flow field (code~\ref{lst:IndDispCode}). The combination of these two model elements represents a previously trained model to predict surface drifter trajectories. While it neglects time-variability and small scales in ocean currents altogether, this model is sufficient to capture the mean pathway of sea water through this region -- the well-known loop current feeding the Gulf Stream via Florida Strait (Fig.~\ref{fig:drifters_speed_simu}).

@@

\begin{lstlisting}[
    language = Julia,
    numbers=left,
    label={lst:IndDispCode},
    caption={Simulate surface drifter pathways in the Gulf of Mexico (Fig.~\ref{fig:drifters_speed}) from flow fields $u,v$ and initial positions $x,y$.}
]
using Drifters

@@

In the {\it digital twins for ocean robots} framework (DTOR), interactivity is facilitated by Julia and its large ecosystem of software packages. In particular, \pkg{Makie.jl} \cite{Makie} and \pkg{Pluto.jl} \cite{Pluto} let us operate even complex modeling workflows from notebooks and apps. The tutorial examples in \pkg{OceanRobots.jl}, \pkg{ArgoData.jl}, \pkg{ClimateModels.jl}, and \pkg{MITgcm.jl} \cite{Forget2024b} are notably available as Pluto notebooks. Interaction between model and data is further enabled by the rich Julia ecosystem for machine learning, data assimilation, parameter inference, and so on. Fig.~\ref{fig:MLP} provides an example where we used \pkg{Flux.jl} \cite{Innes2018} to train a neural network to predict chlorophyll concentration (linked to marine microbe abundance) from environmental parameters as done in the CANYON model \cite{Sauzede2017,Bittig2018}.

The core of DTOR is formed by \pkg{OceanRobots.jl}, \pkg{ArgoData.jl}, and \pkg{MeshArrays.jl} for the physical twins, along with \pkg{ClimateModels.jl}, \pkg{MITgcm.jl}, and \pkg{Drifters.jl} for the virtual twins (i.e., predictive modeling). The \pkg{ClimateModels.jl} interface streamlines the use of models implemented in various languages. Our current model hierarchy supported via \pkg{ClimateModels.jl} includes, but is not limited to, the models used in this paper. DTOR can notably simulate observations in the future based on climate model predictions using \pkg{ClimateModels.jl}, which provides two options for this kind of applications -- either querying the CMIP6 archive of model output \cite{CMIP6} or using a fast emulator such as Hector \cite{Hector2015} (Fig.~\ref{fig:globalwarming}).

Within our model hierarchy, future climate scenarios like Fig.~\ref{fig:globalwarming} can be downscaled using gridded climatologies and reanalyses \cite{Forget2010,Forget2015a,Forget2024a}. The ECCO4 and OCCA2 reanalyses are simple to rerun with perturbed surface forcing fields via \pkg{MITgcm.jl}, which is partiucularly convenient for such applications. Multi-decadal solutions produced in this way can then be further downscaled using km-scale model output (Fig.~\ref{fig:km_scale}). At the end of this modeling workflow, \pkg{OceanRobots.jl}, \pkg{MeshArrays.jl} and \pkg{Drifters.jl} enable calculations such as Figs.~\ref{fig:Argo_Float_simu} and \ref{fig:drifters_speed_simu}, for any of our models. DTOR can also take advantage of gridded satellite data, incl. sea surface temperature and sea level anomalies, and it is sometimes possible to use these instead of ocean reanalyses.

@@

Further development of the DTOR framework is expected to proceed in several directions. First, we'd like to integrate additional types of ocean observing platforms, starting with those that do not have a data structure listed yet in Tab.~\ref{tab:robots}. Observational networks that focus on the coastal ocean, a specific region, or local field experiments could be supported in the future. A top level API to interact with DTOR as a whole through cloud services is also envisioned.

@@

Our current focus is on geospatial analyses that track global warming and marine heatwaves using Argo data \cite{Forget2024a}. Another example from our research is with the C-Streams research program (UK-US) which is releasing drifters near Florida Strait to track the nutrient stream that is associated with the Gulf Stream. The example in Fig.~\ref{fig:Cstream} is based on the monthly mean ECCO climatology for ocean transports \cite{Forget2015a,Forget2019,Rousselet2021}, and using \pkg{Climatology.jl}, \pkg{MeshArrays.jl}, and \pkg{Drifters.jl} to calculate pathways. Other applications at the scale of oceanic basins include tracking ocean plastic pollution, monitoring biological impacts of extreme events and global warming, or the optimization of the global climate monitoring fleet of ocean robots for e.g. cost. More locally, DTOR could be used in the context of field experiments such as S-MODE, to choose an example from the recent past. Related projects on digital twins, not focused on exploiting Julia, include the EU funded {\it Destination Earth} program, Mercator Ocean International's {\it Digital Twin Ocean}, and the UN Decade Action's DITTO initiative ({\it Digital Twins of The Ocean}).

>>>>

DTOR is a framework to access and simulate the global fleet of ocean observing devices (or {\it ocean robots}) that monitor climate change. It brings these observations to Julia and let's us pair ocean robots with virtual counterparts (or {\it twins}). Digital twins provide a bridge to predictive models that enables machine learning. In turn, simulating observations in a digital environment can help evaluate observational strategies a priori, during deployment, or afterwards. In this paper we present the framework, review simulated ocean robots, and discuss envisioned applications.

@@

The DTOR framework was introduced at the Symposium on Advances in Ocean Observation in 2022, and later presented to US-CLIVAR and JuliaCon in 2023. The primary goal of DTOR is to associate ocean observing systems (or {\it ocean robots}) with virtual counterparts (or {\it twins}), and do it for the whole fleet of ocean robots that are currently at sea or have observed the Ocean in the past (Fig.~\ref{fig:ocean-ops}). The virtual twins are to be created through numerical model simulations, which can come in different flavors and languages. The simulation of observing systems in the future, as climate change progresses, is also part of the scope of the DTOR project. At JuliaCon 2023, we presented a solution to (1) access and manipulate the data collected by ocean robots (inc. \pkg{OceanRobots.jl} and \pkg{ArgoData.jl}), and (2) simulate such observations in a digital environment using a hierarchy of models (incl. \pkg{ClimateModels.jl} and \pkg{MITgcm.jl}).

The model hierarchy includes fast climate model emulators, global model output, ocean reanalyses, high-resolution models, and several ways to represent marine ecosystems. Through a streamlined workflow, \pkg{ClimateModels.jl} (JuliaCon 2021, 2023) makes it easy to operate these models that can provide a digital environment for the virtual ocean robot fleet to observe and navigate. \pkg{Drifters.jl} \cite{Forget2021} can be use to predict pathways of ocean robots that tend to follow ocean currents. \pkg{MeshArrays.jl} adds basic geospatial support for global climate model grids, providing capabilities such as interpolation and geolocation on a grid. The framework leverages and links to a variety of highly capable Julia packages from the community. It notably provides extension for visualization with \pkg{Makie.jl}, and tutorial examples in the form of \pkg{Pluto.jl} notebooks.

@@

Part of the fleet of ocean robots currently at sea is depicted in Fig.~\ref{fig:ocean-ops}, focusing on some of the most common observing platforms. To create this map, \pkg{OceanRobots.jl} queries the meta-data-base from Ocean-OPS.org, which keeps track of the whole ocean fleet of scientific observation platforms in real time, and provides a restful API. The \pkg{OceanRobots.jl} package provides a simple interface to this API via the `OceanOPS` module. \pkg{OceanRobots.jl} defines data structures to access and utilize data from common ocean robot types such as  `SurfaceDrifter` and `ArgoFloat` (shown in Fig.~\ref{fig:ocean-ops}). A non-exhaustive list of ocean observing platform categories within the scope of \pkg{OceanRobots.jl} is provided in Tab.~\ref{tab:robots}.

@@

Argo profiling floats (blue dots in Fig.~\ref{fig:ocean-ops}; Code~\ref{lst:ArgoCode}) are one of our main tools to monitor global warming below the sea surface. These devices provide a less detailed view of oceanic pathways (Fig.~\ref{fig:Argo_Float}, top panels) than surface drifters do (e.g., Fig.~\ref{fig:drifters}), since Argo floats only report their location once every ten days. However, Argo floats have a crucial diving capability that surface drifters don't have -- they go up and down the water column to measure temperature and salinity (T,S; bottom panels of Fig.~\ref{fig:Argo_Float}). As a result, the advent of the Argo array in the early 2000s \cite{Roemmich2019} opened up a whole new era of geospatial analysis, state estimation, and parameter inference over the global Ocean  \cite{Forget2007,Forget2010,MFF14,Forget2015a,FFL15,FP15,Forget2024a}. There were 3837 Argo floats at sea on `2024/05/27`, and a total of 18730 in the Argo data base. Since the Argo array is such an important observing system, a dedicated package called \pkg{ArgoData.jl} was created that \pkg{OceanRobots.jl} uses under the hood.

@@

\pkg{OceanRobots.jl} brings these key climate data sets to the Julia community. It provides a bridge to climate scientists working on observations and models, who are interested in leveraging the powerful Julia software ecosystem (e.g., for numerical modeling, machine learning, and statistical analysis), and have much expertise to contribute. The development of \pkg{OceanRobots.jl} aims to help advance (1) how we understand and simulate observations that monitor climate change in the oceans, and (2) climate literacy and education by providing simple apps that anyone can use.

@@

Our ability to simulate the data produced by ocean robots is directly linked to our ability, or lack thereof, to explain observed variations and decipher mechanisms from data. Numerical modeling is thus often motivated and driven by observations. Many research activities involve combining ocean models and observations, with the typical goal of learning model parameters and dynamics from data \cite{Forget2010,Forget2015a,FFL15,FP15,Forget2024a}. Observing system simulations in a digital environment (using models) enable a wide range of applications -- to test out innovative platform or sensor designs ahead of deployment, to optimize global monitoring strategies, or to guide deployed assets in real time for example.

The two simple examples presented below illustrate the simulation of (1) environmental sensors and (2) observing platforms. First, in Fig.~\ref{fig:Argo_Float_simu} we simulate an Argo data collection for temperature (T) and salinity (S) by sampling an ocean climatology along the track of the Argo float (the one from Fig.~\ref{fig:Argo_Float}). This calculation requires (1) a model prediction of $T,S(x,y,z,t)$, which can be based on statistical or mechanistic models on a grid, and (2) tools that deal with the Earth geography, and can localize observations on a grid, and perform interpolation tasks.

@@

Our second example focuses on the simulation of observing platforms moving with ocean currents. This basic simulation of surface drifter pathways uses the geostatistical average of velocities calculated in Fig.~\ref{fig:drifters_ve} as a predictive model for the $U,V(x,y,t)$ velocity fields. This gridded data set is simply provided as input to \pkg{Drifters.jl}, which then releases virtual drifters and calculates their trajectories following the flow field (code~\ref{lst:IndDispCode}). The combination of these two model elements represents a previously trained model to predict surface drifter trajectories. While this simple calculation neglects time-variability and small scales in ocean currents altogether, the model is sufficient to capture the mean pathway of sea water through this region -- the well-known loop current feeding the Gulf Stream via Florida Strait (Fig.~\ref{fig:drifters_speed_simu}).

@@

\begin{lstlisting}[
    language = Julia,
    numbers=left,
    label={lst:IndDispCode},
    caption={Simulate surface drifter pathways in the Gulf of Mexico (Fig.~\ref{fig:drifters_speed}) from flow fields $u,v$ and initial positions $x,y$.}
]
using Drifters, DataFrames

@@

In the {\it digital twins for ocean robots} framework (DTOR), interactivity is facilitated by Julia and its large ecosystem of software packages. In particular, \pkg{Makie.jl} \cite{Makie} and \pkg{Pluto.jl} \cite{Pluto} let us operate even complex modeling workflows from notebooks and apps. The tutorial examples in \pkg{OceanRobots.jl}, \pkg{ArgoData.jl}, \pkg{ClimateModels.jl}, and \pkg{MITgcm.jl} \cite{Forget2024b} are notably available as Pluto notebooks. Interaction between model and data is further enabled by the rich Julia ecosystem for machine learning, data assimilation, parameter inference, etc. Fig.~\ref{fig:MLP} provides an example where \pkg{Flux.jl} \cite{Innes2018} is used to train neural networks to predict chlorophyll concentration (linked to marine microbe abundance) from environmental parameters as done in the CANYON model \cite{Sauzede2017,Bittig2018}.

The core of DTOR is formed by \pkg{OceanRobots.jl}, \pkg{ArgoData.jl}, and \pkg{MeshArrays.jl} for the physical twins, along with \pkg{Climatology.jl}, \pkg{ClimateModels.jl}, \pkg{MITgcm.jl}, and \pkg{Drifters.jl} for the virtual twins (i.e., predictive modeling). The \pkg{ClimateModels.jl} interface streamlines the use of models implemented in various languages. The model hierarchy already includes the models used in this paper, and is easy to extend via pkg{ClimateModels.jl}. DTOR can notably simulate observations in the future based on climate model predictions, and \pkg{ClimateModels.jl} provides two options for this kind of applications -- either querying the CMIP6 archive of model output \cite{CMIP6} or using a fast emulator such as Hector \cite{Hector2015} (Fig.~\ref{fig:globalwarming}).

Within our model hierarchy, future climate scenarios like Fig.~\ref{fig:globalwarming} can be downscaled using gridded climatologies and reanalyses \cite{Forget2010,Forget2015a,Forget2024a}. The ECCO4 and OCCA2 reanalyses are simple to rerun with perturbed surface forcing fields via \pkg{MITgcm.jl}, which is partiucularly convenient for such applications. Multi-decadal solutions produced in this way can then be further downscaled using km-scale model output (Fig.~\ref{fig:km_scale}). At the end of this modeling workflow, \pkg{OceanRobots.jl}, \pkg{MeshArrays.jl} and \pkg{Drifters.jl} enable calculations such as Figs.~\ref{fig:Argo_Float_simu} and \ref{fig:drifters_speed_simu}, for any of our models. DTOR can also take advantage of gridded satellite data, incl. sea surface temperature and sea level anomalies via \pkg{Climatology.jl}, and it is sometimes possible to use these instead of ocean reanalyses.

@@

Further development of the DTOR framework is expected to proceed in several directions. First, we'd like to integrate additional types of ocean observing platforms, starting with those that do not have a data structure listed yet in Tab.~\ref{tab:robots}. Observational networks that focus on either the coastal ocean, a specific region, or local field experiments could be supported in the future. A top level API to interact with DTOR as a whole through cloud services is also envisioned.

@@

Our current focus is on geospatial analyses that track global warming and marine heatwaves using Argo data \cite{Forget2024a}. Another example from our research is with the C-Streams observational program (UK-US) which is releasing drifters near Florida Strait to track the nutrient stream that is associated with the Gulf Stream. A simulation of these pathways is shown in Fig.~\ref{fig:Cstream}, which is based on the monthly mean ECCO4 climatology for ocean transports \cite{Forget2015a,Forget2019,Rousselet2021}, and uses \pkg{Climatology.jl}, \pkg{MeshArrays.jl}, and \pkg{Drifters.jl} to calculate pathways. Other applications at the scale of oceanic basins include tracking ocean plastic pollution, monitoring biological impacts of extreme events and global warming, or the optimization of the global climate monitoring fleet of ocean robots. DTOR could be used to pilot field experiments (e.g., S-MODE, to choose an example from the recent past). Related projects on digital twins, not focused on exploiting Julia but interested in ocean observations, include the EU funded {\it Destination Earth} program, Mercator Ocean International's {\it Digital Twin Ocean}, and the UN Decade Action's DITTO initiative ({\it Digital Twins of The Ocean}).
